Efficient, Robust and Exact Boolean Operations on N Triangular Mesh Primitives

### 2016-09-27 summary.tex
9.27

We develop an unconditionally robust and exact boolean operation method for triangle meshes without using exact arithmetic. The robustness of our method is based on the observations by Sugihara and Iri \cite{sugihara1990solid} that boolean operations can be performed requiring geometry constructions if representations are based on planes and not vertices. This idea of using plane-based representation has been developed by Berstein and Fussell \cite{bernstein2009fast} by combining it with binary search partition (BSP) techniques \cite{naylor1990merging,thibault1987set}. Later, Campen et. al. \cite{campen2010exact} improved the method using octrees to resolve intersections locally, avoiding inefficient large BSP tree construction. Despite of that, Campen et. al.'s method still suffers from performance problem because of the complex conversions among B-reps, plane-based representations and BSP representations. Also, the incoherence between BSP representation, where boolean operations actually happens, and B-reps, which are input and output meshes, requires extra steps to connect the reconstructed intersection areas with the untouched meshes, leading to extra computation burden and complex topology.

We adopt plane-based techniques and strictly follow the principle of no geometry construction, yet develop a very different framework based on halfedges, requiring only one-time conversion to the plane representation in intersection areas. Since the geometry connectivity is stored in halfedge, it allow us to take advantage of geometry connectivity for classification acceleration \cite{feito2013fast,ogayar2015deferred,updegrove2016boolean}, which is hard to for a BSP-based method. In addition, our method can perform boolean operations with more than two primitives in one call, instead of recursively doing binary boolean operations. In this way, extra complexity is introduced. But we can save much computation time by avoiding repetitive computation. The major difficulty of applying plane-based representation on halfedge is by the incoherence between the two representations. Halfedge is a vertex-based representation. Many geometry operations (e.g., coordinates projection and compute the barycenter coordinate of triangle) available on halfedge can be very complex and thus inefficient under plane-based representation. On the other hand, we have to compute under plane-based representation for exactness and robustness. Therefore, we manage to limit our computation of boolean operations to several basic plane-based predicates. The experiment shows that our method is much faster than Campen et. al.'s method, while only about two to three times slower than the state-of-art non-robust method \cite{douze2015quickcsg}.

@@

The concept of plane-based representation of polygonal meshes was first described by Sugihara and Iri \cite{sugihara1990solid}. Such representation provides important advantage that for boolean operations, no new primary geometry information has to be constructed to obtain the result polyhedron---it is composed of a subset of the planes of the input polyhedra. Berstein and Fussell \cite{bernstein2009fast} combined the plane-based representation with binary space partitioning (BSP) \cite{thibault1987set,naylor1990merging} to develop an unconditionally robust boolean operation method. Compared to volume-grid based methods, their BSP-based method increases mesh complexity more moderately and the input topology can be reconstructed better. However, it requires conversions between input data structures and plane-based representation which are inefficient from both storage and time views. Also, the merge of two BSP trees is an $O(mn)$ time complexity process, where $m$ and $n$ are the size of the trees. It makes this method impractical for very large input meshes. Campen and Kobbelt \cite{campen2010exact} improved their method by localization using an octree. And the refinement only takes place locally near intersections.

@@

For efficiency, the exactness and robustness should not rely on arbitrary precision arithmetic. Instead, we choose the plane-based techniques from \cite{campen2010exact}. However, as we discussed in section \ref{sec:pbrelated}, BSP structure does not support geometry connectivity query in nature, and we do not want fussy conversions between different representations. Therefore, we use halfedge instead as our major data structure and embed plane-based representation into it.

@@

However, naive halfedge is far from enough. The vertex coordinates in halfedge are usually represented by float point number, which cannot exactly represent the newly introduced intersection points. Therefore, we utilize plane-based representation to avoid computing new coordinates. In addition, we detect all coincident vertices among input meshes (including intersection vertices). This information is used for avoiding repetitive vertices in the final mesh, reconstructing geometry connectivity, and ensuring the topology correctness.

@@

This step is to detect intersections between primitives in a robust and exact way. Our algorithm is largely based on vertex-based M\"{o}ller's algorithm \cite{moller1997fast}, which is very efficient to compute intersection between two triangles. However, naive implementation of M\"{o}ller's algorithm always incurs robustness issues. To avoid an unexpected failure caused by numeric errors, we integrate plane-based representation of polyhedra into M\"{o}ller's algorithm. All intersection vertices are represented using plane intersections to avoid geometry constructions. Also, we carefully identify all degenerate cases of triangle intersections, including point intersection, line intersection and coplanar intersection, and respectively discuss how to deal with each of them. Details are provided in Section \ref{section:isect}.

@@

Once all intersections are detected, we need to tessellate the input primitives and construct the Linked Halfedges. In many methods like \cite{ogayar2015deferred}, constraint Delaunay triangulation (CDT) is applied to perform tessellation, since the intersections are naturally the constraints of CDT. However, for a CSG with more than two primitives, the intersection vertices may be generated by intersections of three faces from different primitives, which cannot be computed during by triangle-triangle intersections in the previous step. Intersection line segments may overlap or intersect with each other, and cannot be used for constraints of CDT directly. In addition, as our intersection vertices are represented exactly by planes, implementation of CDT are complex and inefficient, as most CDT methods were not designed to handle planes and requires explicit coordinates. Therefore, we design a structure suited for plane-based representation, called \emph{Tess-graph}, to guide the tessellation of each single face, and develop a very fast but exact method to construct Linked Halfedges from Tess-graphs. Details are shown in section \ref{sec:tessellation}.

@@

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{p-reps}\\
  \caption{Plane-based representation of triangles. The yellow triangle is represented by three bounding planes (green) and the supporting plane where the triangle lies.}\label{fig:p-reps}
\end{figure}

As we discussed in Section \ref{sec:paradigm}, one of the problem in boolean operations under vertex-based representation is it inevitably introduces new vertices, whose coordinates cannot be exactly represented using fix-precision float-point number. On the other hand, if plane-based representation is used, no new geometry information needs to be constructed. Therefore, we can use only geometric decision predicates to perform boolean operations. As the precision of input coordinates is finite, we can determine in advance the upper boundary of precision needed to perform a precise decision, with which we can use fixed precision predicates techniques to gain much performance while keeping exactness. For efficiency, we implement the predicates using filtering techniques proposed by Shewchuk \cite{shewchuk1997adaptive}.

Using plane-based representation, each face $f$ of polyhedron is represented by a supporting plane $\bm{p}_{f,s}$, where the face lies, and a list bounding planes $\{\bm{p}_{f,b}^i \ \vert\  i = 1, 2, 3\}$, each has a corresponding edge (Fig. \ref{fig:p-reps}). Edge line is represented by intersection of $\bm{p}_{f,s}$ and the related $\bm{p}_{f,b}^i$. Corner vertex is the intersection of $\bm{p}_{f,s}$ and two consecutive bounding planes $\bm{p}_{f,b}^i$ and $\bm{p}_{f,b}^{i+1}$.

We also introduce some commonly used notations for better presentation in the rest of paper. A plane $\bm{p}$ can be represented by four scalar coefficients $(a, b, c, d)$. The normal of $\bm{p}$ is denoted as $\bm{n}(\bm{p})$. A line $\bm{l}$ can be represented by intersection of two planes $(\bm{p}_l^0 \cap \bm{p}_l^1)$, or in short, $\bm{l}\colon(\bm{p}_l^0 \cap \bm{p}_l^1)$. The positive direction of the line $\bm{l}$ is defined by $\bm{n}(\bm{p}_l^0) \times \bm{n}(\bm{p}_l^1)$. A point $\bm{v}$ can be represented by non-trivial plane triples $(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$, or in short, $\bm{v}\colon(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$.

@@

We use Campen et al.'s method \cite{campen2010exact} for exact conversion of triangles (represented by three vertices) from vertex-based representation to plane-based representation. Given that the input vertex coordinate has $L$-bit precision including sign, the precisions of the four plane coefficients are $L_a, L_b, L_c, L_d$, which have to meet the following requirements:
\begin{equation}
\begin{split}
&L_a=L_b=L_c\ge 2(K-1)+1+1,\\
&L_d\ge(L_a-1)+(L-1)+2+1.
\end{split}
\end{equation}
Here $K$ is the maximum bits required to represent edge vector components in input meshes. If we use $M$ bits to store each plane coefficient, basically $M$ has to be greater than $L_d$. In our implementation, we actually use a more strong constraint $M \ge L_d+1$. Under this constraint, given a $L$-bit vertex $\bm{v}$, we can compute the orientation of plane with respect to $\bm{v}$ by the sign $\bm{n}\cdot\bm{v} + d$, which can be exactly computed using common float point arithmetic with limited $M$-bit precision. It allows very fast plane orientation predicates which benefits much the early rejection in our intersection computation ($\S$ \ref{section:isect}).

@@

\section{Exact Intersection Detection}

\label{section:isect}
In this step, intersections between faces are computed through the triangle-triangle intersection test within each critical cell of octree. We adopt M\"{o}ller's intersection detection algorithm \cite{moller1997fast} on account of its efficiency and simplicity. However, because collision detection algorithms are often accompanied by non-robustness problems, a naive implementation may cause unpredictable results when there comes degenerate cases. Therefore, we integrate plane-based geometry representation into M\"{o}ller's algorithm to make it unconditionally robust and exact. In the following, we first make a quick review of M\"{o}ller's algorithm, then discuss the way to integrate plane-based geometry. At last, we talk about how to deal with degenerate case, such as intersecting on points and coplanar case.

\subsection{Vertex-Based Method Review}

@@

M\"{o}ller's algorithm computes the intersection between two triangles $t_1$ and $t_2$ in three steps (Fig. \ref{fig_projection}).
\begin{itemize}[leftmargin=0.45cm]
\item[1)] An early rejection is performed by testing whether $t_1$ intersects $p^s(t_2)$, the supporting plane of $t_2$. This is a necessary condition of intersection between $t_1$ and $t_2$. If this test is passed, then the same is performed for $t_2$ and $p^s(t_1)$.
\item[2)]The intersection between $t_1$ and $p^s(t_2)$, denoted as $Seg_1$, and the intersection between $t_2$ and $p^s(t_1)$, denoted as $Seg_2$, are separately computed .
 \item[3)]The intersection between $t_1$ and $t_2$ is determined by computing the overlap area of $Seg_1$ and $Seg_2$ .
\end{itemize}

Conventional implementations of M\"{o}ller's algorithm use vertex-based geometry, which is neither exact nor robust.  Although implementing M\"{o}ller's algorithm with arbitrary precision arithmetic could make it robust, it is too costly for a large CSG evaluation. Actually, the non-robustness of this algorithm is from geometry \emph{constructions} \cite{bernstein2009fast}, which compute new coordinates of geometry objects based on the known coordinates of existing ones. In M\"{o}ller's algorithm, the coordinates of $Seg_1$ and $Seg_2$ (Fig. \ref{fig_projection}) are computed as intermediate results. We use plane-based geometry to solve these problems by avoiding explicitly computing the coordinates of new vertices.

@@

\label{sec:embed}
According to the method described in section \ref{sec:convert}, triangle faces $f$ are firstly converted to its plane-based representation: a supporting plane $p^s(f)$ surrounded by three bounding planes $\{p^b_i\ |\  i = 1,2,3\}$. We integrate the plane-based geometry into each of the tree steps of M\"{o}ller's algorithm as following:

@@

\begin{itemize}[leftmargin=0.45cm]
\item[1)] The basic routine of the first step is given two faces $t_1$ and $t_2$, computing the orientation of the supporting plane (denoted as $p(t_1)\colon(a, b, c, d)$) of a face with respect to a certain vertex (denoted as  $v_i(t_2)\colon(x, y, z)$) of the other face. That is to compute:
\begin{equation}\label{eq:plane}
sign(ax+by+cz+d).
\end{equation}
Reviewing the discussion of section \ref{sec:convert},  the coordinates of vertices are represented by three L-bit float point numbers.Precisions of the four coefficients of the supporting plane are $(P^\star, P^\star, P^\star, P)$. Here $P$ is defined in equation \ref{eq:precision}, and $$P^\star=P-(L-1)-2.$$ Using the enforced criterion $M \ge P+1$, we find that $M$ bits are enough to perform the predicate of equation \ref{eq:plane}. This allow us to perform the early discard efficiently.
\item[2)]In step 2, the end points of $Seg_1$ and $Seg_2$ are implicitly represented using intersections of plane triples. Take $Seg_1$ as an example. As illustrated in Fig. \ref{fig:isect}, $Seg_1$ are the intersection between $t_1$ and $p_s(t_2)$. Because the edge $e_i(t_1)$ can be represented as $(p^s( t_1), p^b_i(t_1))$, the end points of $Seg_1$ can be represented in the form of $(p^s( t_1), p^s(t_2), p^b_i(t_1))$. In Fig. \ref{fig:isect}, we list all intersection situations between $t_1$ and $p_s(t_2)$, and the corresponding bounding plane $p^b_i(t_1)$ in each situation. The conditions that $t_1$ is inside the plane $p_s(t_2)$ (coplanar cases) will be discussed later in section \ref{sec:degenerate}.
 \item[3)] The intersection between $t_1$ and $t_2$ is the overlap area of $Seg_1$ and $Seg_2$. It can be easily computed by comparing the end points of $Seg_1$ and $Seg_2$ along $l$ (Fig. \ref{fig_projection}), where $l$ is the intersection between $p_s(t_1)$ and $p_s(t_2)$. Because end points are all represented using plane triples, we can use the two-line-one-plane comparison discussed in section \ref{sec:substrates} to perform the comparison.
\end{itemize}

\subsection{Intersection Representation}
\label{sec:ir}

// change the title to Indentifying intersections

The output of this stage, or what we call intersection record (IR), is the data structure which organize the results of triangle-triangle intersection test. IR is also the input of tessellation, and is in deep couple with the tessellation method. It is hard to give full idea of why the IR is like that before we introduce our tessellation algorithm. Therefore, in this section, we first give several requirements for IR from tessellation stage, and then discuss the IR structure. To have a comprehensive understanding of IR, readers may need to review this section after finish reading section \ref{sec:tessellation} about tessellation stage.

We recall our definition of Linked Halfedge in section \ref{sec:meshes}, we need to transfer necessary information by IR to tessellation stage for Linked Halfedge construction. For each intersection line segment, we use a five-component vector {\color{red}{$\Gamma \colon(t, v_0, v_1, p, C_{P_h \backslash P_i})$}} to describe it. Here, $v_b$ and $v_e$ are the two end points of $\Gamma$. Component $f$ is the face where the $\Gamma$ lies. Component $C$ is the intersection context, and will be further explained later. Component $p$ is the plane representation of the line where $\Gamma$ lies (denoted as $\vec{n}_\Gamma$). Normally, line is represented by two planes. However, one of the planes for $\vec{n}_\Gamma$ must be $f$'s supporting plane $p^s(f)$, so we just omit it, and use the other plane as the representation for $\vec{n}_{\boldsymbol{\Gamma}}$ .

To illustrate, assume two triangle faces $t_1$ and $t_2$, from primitive $P_1$ and $P_2$, intersect at line segment ${\boldsymbol{\Gamma}}$. Note that there are two versions of ${\boldsymbol{\Gamma}}$ for each of the two faces respectively, denoted as ${\boldsymbol{\Gamma}}(t_1)$ and ${\boldsymbol{\Gamma}}(t_2)$. For ${\boldsymbol{\Gamma}}(t_1)$, $v_b$ and $v_e$ are the two end points of ${\boldsymbol{\Gamma}}$ and $f=t_1$. The representation plane $p$ is $p^s(t_2)$. Note that $p^s(t_2)$ must not be coplanar with $t_1$. The intersection context $C=t_2$ in most cases. It means that, in primitive $P_2$, ${\boldsymbol{\Gamma}}(t_1)$ lies on the face $t_2$. However, sometimes, intersection may lie on the edge, instead of face,  of other primitive. In that case,  the intersection context is an edge, and we discuss such degenerate case in the next section. The situation of ${\boldsymbol{\Gamma}}(t_2)$ is similar we omit the discussion here.

In the Linked Halfedge, there must not be any repetitive vertices in order to reconstruct the correct topology. Therefore, repetitive vertices elimination must be performed. Checking whether there are coincident points in a global scope can be quite inefficient. Our implementation use a localized scheme that store the reference of intersection vertices into their corresponding geometry location, like edges and faces. Therefore, coincidence detection performed within the geometry elements and repetitive vertices elimination can be very fast.

\subsection{Degenerate Cases}
\label{sec:degenerate}

Assume there are two faces $t_1$ and $t_2$ from primitive $P_1$ and $P_2$ respectively. These two faces intersect on line segment ${\boldsymbol{\Gamma}}$, which has two versions of IR ${\boldsymbol{\Gamma}}(t_1)$ and ${\boldsymbol{\Gamma}}(t_2)$ for the two faces respectively. Without loss of generality, we only discuss the configuration of ${\boldsymbol{\Gamma}}(t_1)$, as ${\boldsymbol{\Gamma}}(t_2)$ can be figured out similarly. We classify all degenerate cases into three categories by the dimension of intersection. Each of them is discussed separately in the following.

Before we go into details, there are three principles that we follow to solve degenerate cases:
\begin{itemize}
\item We want our method be unconditional robust and exact, so all degenerate cases should be well handled (completeness).
  \item To control the complexity of tessellation, we want degenerate cases to be invisible to tessellation, that is, degenerate cases have the same format of output, in our method, IR, as non-degenerate cases (invisibility).
  \item we want the outcome of tessellation be as simple as possible. That is, intersection line segments that will not help tessellation, if recognized, are disposed (concision).
\end{itemize}
The criterion of whether an intersection help tessellation is by its necessity: if the intersection is not presented, whether some faces from the linked halfedge will cross the boundary of primitives. The word 'cross' do not only include the situation that a face is partly outside and partly inside of a primitive, but also can mean that a face is partly inside (outside) and part on the boundary of the primitive.

\subsubsection{Intersects on point}
\label{sec:ipoint}
When two triangles intersect on a single point, there are two different conditions (Fig. \ref{fig:isolated}), depending on whether the point is isolated. If the point is isolated, it means that there is no other intersection line segment sharing the point. This condition may cause non-manifold surface because of the singularity of the isolated intersection point. Also, isolated intersection point will not affect the tessellation of the intersection face. On the other hand, if the intersection point is not isolated, there must be other intersection line segments that contain the point, thus the point will be introduced at least twice. Therefore, no matter which condition it is, when two triangle faces intersect on point, we can just omit this intersection and this will not affect the correctness of the final topology.

@@

\subsubsection{Intersects on face edges}

// intersect on face edges does not create new line segments, it is part of the edges from other meshes, and will to be duplicated.

// whose edge? need to be further explained and better defined. an edge intersection divide edge in a triangle and will be duplicated in the other

>>>>

We develop an unconditionally robust and exact boolean operation method for triangle meshes without using exact arithmetic. The robustness of our method is based on the observations by Sugihara and Iri \cite{sugihara1990solid} that boolean operations can be performed requiring geometry constructions if representations are based on planes and not vertices. This idea of using plane-based representation (P-rep) has been developed by Berstein and Fussell \cite{bernstein2009fast} by combining it with binary search partition (BSP) techniques \cite{naylor1990merging,thibault1987set}. Later, Campen et. al. \cite{campen2010exact} improved the method using octrees to resolve intersections locally, avoiding inefficient large BSP tree construction. Despite of that, Campen et. al.'s method still suffers from performance problem because of the complex conversions among B-reps, P-reps and BSP representations. Also, the incoherence between BSP representation, where boolean operations actually happens, and B-reps, which are input and output meshes, requires extra steps to connect the reconstructed intersection areas with the untouched meshes, leading to extra computation burden and complex topology.

We adopt plane-based techniques and strictly follow the principle of no geometry construction, yet develop a very different framework based on halfedges, requiring only one-time conversion to the plane representation in intersection areas. Since the geometry connectivity is stored in halfedge, it allow us to take advantage of geometry connectivity for classification acceleration \cite{feito2013fast,ogayar2015deferred,updegrove2016boolean}, which is hard to for a BSP-based method. In addition, our method can perform boolean operations with more than two primitives in one call, instead of recursively doing binary boolean operations. In this way, extra complexity is introduced. But we can save much computation time by avoiding repetitive computation. The major difficulty of applying P-rep on halfedge is by the incoherence between the two representations. Halfedge is a vertex-based representation. Many geometry operations (e.g., coordinates projection and compute the barycenter coordinate of triangle) available on halfedge can be very complex and thus inefficient under P-rep. On the other hand, we have to compute under P-rep for exactness and robustness. Therefore, we manage to limit our computation of boolean operations to several basic plane-based predicates. The experiment shows that our method is much faster than Campen et. al.'s method, while only about two to three times slower than the state-of-art non-robust method \cite{douze2015quickcsg}.

@@

The concept of P-rep of polygonal meshes was first described by Sugihara and Iri \cite{sugihara1990solid}. Such representation provides important advantage that for boolean operations, no new primary geometry information has to be constructed to obtain the result polyhedron---it is composed of a subset of the planes of the input polyhedra. Berstein and Fussell \cite{bernstein2009fast} combined the P-rep with binary space partitioning (BSP) \cite{thibault1987set,naylor1990merging} to develop an unconditionally robust boolean operation method. Compared to volume-grid based methods, their BSP-based method increases mesh complexity more moderately and the input topology can be reconstructed better. However, it requires conversions between input data structures and P-rep which are inefficient from both storage and time views. Also, the merge of two BSP trees is an $O(mn)$ time complexity process, where $m$ and $n$ are the size of the trees. It makes this method impractical for very large input meshes. Campen and Kobbelt \cite{campen2010exact} improved their method by localization using an octree. And the refinement only takes place locally near intersections.

@@

For efficiency, the exactness and robustness should not rely on arbitrary precision arithmetic. Instead, we choose the plane-based techniques from \cite{campen2010exact}. However, as we discussed in section \ref{sec:pbrelated}, BSP structure does not support geometry connectivity query in nature, and we do not want fussy conversions between different representations. Therefore, we use halfedge instead as our major data structure and embed P-rep into it.

@@

However, naive halfedge is far from enough. The vertex coordinates in halfedge are usually represented by float point number, which cannot exactly represent the newly introduced intersection points. Therefore, we utilize P-rep to avoid computing new coordinates. In addition, we detect all coincident vertices among input meshes (including intersection vertices). This information is used for avoiding repetitive vertices in the final mesh, reconstructing geometry connectivity, and ensuring the topology correctness.

@@

This step is to detect intersections between primitives in a robust and exact way. Our algorithm is largely based on vertex-based M\"{o}ller's algorithm \cite{moller1997fast}, which is very efficient to compute intersection between two triangles. However, naive implementation of M\"{o}ller's algorithm always incurs robustness issues. To avoid an unexpected failure caused by numeric errors, we integrate P-rep of polyhedra into M\"{o}ller's algorithm. All intersection vertices are represented using plane intersections to avoid geometry constructions. Also, we carefully identify all degenerate cases of triangle intersections, including point intersection, line intersection and coplanar intersection, and respectively discuss how to deal with each of them. Details are provided in Section \ref{section:isect}.

@@

Once all intersections are detected, we need to tessellate the input primitives and construct the Linked Halfedges. In many methods like \cite{ogayar2015deferred}, constraint Delaunay triangulation (CDT) is applied to perform tessellation, since the intersections are naturally the constraints of CDT. However, for a CSG with more than two primitives, the intersection vertices may be generated by intersections of three faces from different primitives, which cannot be computed during by triangle-triangle intersections in the previous step. Intersection line segments may overlap or intersect with each other, and cannot be used for constraints of CDT directly. In addition, as our intersection vertices are represented exactly by planes, implementation of CDT are complex and inefficient, as most CDT methods were not designed to handle planes and requires explicit coordinates. Therefore, we design a structure suited for P-rep, called \emph{Tess-graph}, to guide the tessellation of each single face, and develop a very fast but exact method to construct Linked Halfedges from Tess-graphs. Details are shown in section \ref{sec:tessellation}.

@@

\begin{figure}
  \centering
  % Requires \usepackage{graphicx}
  \includegraphics[width=1.7in]{p-reps}\\
  \caption{P-rep of triangles. The yellow triangle is represented by three bounding planes (green) and the supporting plane where the triangle lies.}\label{fig:p-reps}
\end{figure}

As we discussed in Section \ref{sec:paradigm}, one of the problem in boolean operations under vertex-based representation is it inevitably introduces new vertices, whose coordinates cannot be exactly represented using fix-precision float-point number. On the other hand, if P-rep is used, no new geometry information needs to be constructed. Therefore, we can use only geometric decision predicates to perform boolean operations. As the precision of input coordinates is finite, we can determine in advance the upper boundary of precision needed to perform a precise decision, with which we can use fixed precision predicates techniques to gain much performance while keeping exactness. For efficiency, we implement the predicates using filtering techniques proposed by Shewchuk \cite{shewchuk1997adaptive}.

As illustrated in Fig. \ref{fig:p-reps}, using P-rep, each face $f$ with $n$ edges is represented by a supporting plane $\bm{p}_{f,s}$, where the face lies, and a list bounding planes $\{\bm{p}_{f,b}^i \ \vert\  i = 0, 1,...,n-1\}$. Each edge line $\bm{e}_f^i$ is represented by intersection $\bm{p}_{f,s} \cap \bm{p}_{f,b}^i$. Corner vertex $\bm{v}_f^i$ is represented by intersection $\bm{p}_{f,s} \cap \bm{p}_{f,b}^i \cap \bm{p}_{f,b}^{{(i+1)}\bmod{n}}$.

We also introduce some commonly used notations for better presentation in the rest of paper. A plane $\bm{p}$ can be represented by four scalar coefficients $a(\bm{p}), b(\bm{p}), c(\bm{p}), d(\bm{p})$. The normal of $\bm{p}$ is denoted as $\bm{n}(\bm{p})$, which is a normalized vector consisting of the first three plane coefficients. A line $\bm{l}$ can be represented by intersection of two planes $(\bm{p}_l^0 \cap \bm{p}_l^1)$, or in short, $\bm{l}\colon(\bm{p}_l^0 \cap \bm{p}_l^1)$. The positive direction of the line $\bm{l}$ is defined by $\bm{n}(\bm{p}_l^0) \times \bm{n}(\bm{p}_l^1)$. A point $\bm{v}$ can be represented by non-trivial plane triples $(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$, or in short, $\bm{v}\colon(\bm{p}_v^0 \cap \bm{p}_v^1 \cap \bm{p}_v^2)$.

@@

We use Campen et al.'s method \cite{campen2010exact} for exact conversion of triangles (represented by three vertices) from vertex-based representation to P-rep. Given that the input vertex coordinate has $L$-bit precision including sign, the precisions of the four plane coefficients are $L_a, L_b, L_c, L_d$, which have to meet the following requirements:
\begin{equation}
\begin{split}
&L_a=L_b=L_c\ge 2(K-1)+1+1,\\
&L_d\ge(L_a-1)+(L-1)+2+1.
\end{split}
\end{equation}
Here $K$ is the maximum bits required to represent edge vector components in input meshes. If we use $M$ bits to store each plane coefficient, basically $M$ has to be greater than $L_d$. In our implementation, we actually use a more strong constraint $M \ge L_d+1$. Under this constraint, given a $L$-bit vertex $\bm{v_i}$, we can compute the orientation of plane $p_j$ with respect to $\bm{v_i}$ by the sign $\bm{n}(\bm{p_j})\cdot\bm{v_i} + d(\bm{p_j})$, which can be exactly computed using common float point arithmetic with limited $M$-bit precision. It allows very fast plane orientation predicates which benefits much the early rejection in our intersection computation (\S \ref{section:isect}).

@@

\section{Exact Intersection Computation}

\label{section:isect}
In this step, intersections between faces are computed through triangle-triangle intersection test. We adopt M\"{o}ller's algorithm \cite{moller1997fast} on account of its efficiency and simplicity. However, a naive implementation of M\"{o}ller's algorithm may fail to produce correct results because of numerical errors. Therefore, we integrate plane-based geometry representation to make it unconditionally exact and robust. In the following, we first make a quick review of M\"{o}ller's algorithm, and then discuss the way to integrate plane-based geometry. After that, we talk about how to deal with degenerate cases.

\subsection{M\"{o}ller's Vertex-Based Method}

@@

M\"{o}ller's algorithm computes the intersection between two triangles $t_1$ and $t_2$ in three steps as shown in Fig. \ref{fig_projection}:
\begin{itemize}[leftmargin=0.45cm]
\item[1)] An early rejection is performed by testing whether $t_1$ intersects $\bm{p}_{t_2, s}$, the supporting plane of $t_2$. The same test is also done between $t_2$ and $\bm{p}_{t_1, s}$.
\item[2)]The intersection between $t_1$ and $\bm{p}_{t_2, s}$, denoted as $Seg_1$, and the intersection between $t_2$ and $\bm{p}_{t_1, s}$, denoted as $Seg_2$, are separately computed .
 \item[3)]The intersection between $t_1$ and $t_2$ is determined by computing the overlap between $Seg_1$ and $Seg_2$ .
\end{itemize}

The non-robustness of this algorithm is from computing the coordinates of intersection vertices, which is the end points of $Seg1$ and $Seg2$. Although implementation with arbitrary precision arithmetic produces exact coordinates, it is too costly for boolean operations of large CSG. We use plane-based geometry to solve this problem by implicitly representing intersections with planes.

@@

\label{sec:embed}
According to the method described in section \ref{sec:convert}, triangle faces $f$ are firstly converted to its P-rep: a supporting plane $\bm{p}_{t, s}$ surrounded by three bounding planes $\{\bm{p}_{t, b}^i|\ i = 0,1,2\}$. We integrate the plane-based geometry into each of the tree steps of M\"{o}ller's algorithm as following:

@@

\begin{itemize}[leftmargin=0.45cm]
  \item[1)] The basic subroutine of the first step is given two faces $t_1$ and $t_2$, computing the orientation of of a face with respect to vertices from the other face. Reviewing the discussion of \S \ref{sec:convert}, under the constraint of $M \ge P+1$, we can compute the predicate within the precision of $M$ bits. This allow us to perform this early rejection efficiently.
      \vspace{0.5em}
  \item[2)] If $t_1$ and $t_2$ are not coplanar, the end points of $Seg_1$ and $Seg_2$ can be implicitly represented by plane triples. Take $Seg_1$ as an example. $Seg_1$ is the intersection between $t_1$ and $\bm{p}_{t_2, s}$ and the end points of $Seg_1$ are intersections between $\bm{p}_{t_2, s}$ and edges from $t_1$ (denoted as $\bm{e}^i_{t_1}$). Because $\bm{e}^i_{t_1}$ can be represented as $\bm{p}_{t_1, s}\cap \bm{p}^i_{t_1, b}$, the end points of $Seg_1$ can be represented in the form of $\bm{p}_{t_2, s} \cap \bm{p}_{t_1, s} \cap \bm{p}^i_{t_1, b}$. In Fig. \ref{fig:isect}, we list all possible intersecting situations between $t_1$ and $\bm{p}_{t_2, s}$. The two corresponding bounding planes $\bm{p}^i_{t_1, b}$ for each situation are also presented. The coplanar condition will be discussed later in \S \ref{sec:degenerate}.
      \vspace{0.5em}
 \item[3)] The intersection between $t_1$ and $t_2$ is the overlap area of $Seg_1$ and $Seg_2$. It can be easily computed by sorting the end points of $Seg_1$ and $Seg_2$ along the line $\bm{l}\colon \bm{p}_{t_1, s} \cap \bm{p}_{t_2, s}$. Because end points are all represented using plane triples, we can use the linear order of points discussed in \S \ref{sec:substrates} to sort.
\end{itemize}

Each time an intersection is computed, two intersections are generated (one for each triangle face) and the newly introduced vertices are introduced into related meshes. To avoid repetitive vertices in the final result, we do repetition elimination in this stage, merging coincident vertices even among different input meshes. Because we use P-reps, the comparison of vertices is exact.

\subsection{Plane-Based Intersection Representation}
\label{sec:ir}

Once intersections are computed, proper representation is needed to guarantee exact transfer of information and low cost to process. In our method, each line segment intersection $\bm{\Gamma}$ between triangles is represented by our plane-based intersection representation (PBI-rep) as $\{t, \bm{p}_{ext}, (\bm{p}^0, \bm{p}^1),\bm{\Psi}\}$. The first component indicates $\bm{\Gamma}$ is on triangle $t$, and second one $\bm{p}_{ext}$ indicates $\bm{\Gamma}$ lies on a line $\bm{l}_{\bm{\Gamma}}\colon(\bm{p}_{t, s} \cap \bm{p}_{ext})$. The two end points of $\bm{\Gamma}$ are $\bm{l}_{\bm{\Gamma}}\cap\bm{p}^0$ and $\bm{l}_{\bm{\Gamma}}\cap\bm{p}^1$. The last component $\bm{\Psi}$ is a vector that represents the neighborhood of the intersection, which is for fast face classification in later stages. For each component of $\bm{\Psi}$, if $\bm{\Gamma}$ is on the surface of primitive $P_k$, the $k^{th}$ component $\Psi_{k}$ is the edge or face of $P_k$ where $\bm{\Gamma}$ lies. Otherwise, $\Psi_{k}=0$, meaning default empty data.

To illustrate, assume two triangle faces $t_1$ and $t_2$, from primitive $P_a$ and $P_b$, intersect on a line segment. Then it generates two intersections for both triangles respectively, denoted as ${\boldsymbol{\Gamma}}_{12}$ and ${\boldsymbol{\Gamma}}_{21}$. For ${\boldsymbol{\Gamma}}_{12}$, the first two components are $t_1$ and $\bm{p}_{t_2, s}$. The computation of the third component is discussed in \S\ref{sec:embed}. The neighborhood $\bm{\Psi}$ is zero for all components except $\Psi_a$ and $\Psi_b$. In most time, $\Psi_a=t_1$ and $\Psi_b=t_2$. However, sometimes, intersections may lie on the primitive edges instead of faces. In that case, $\Psi_a$ and $\Psi_b$ are edges. We postpone this discussion until \S\ref{sec:degenerate}. The computation of ${\boldsymbol{\Gamma}}_{21}$ is similar.

\subsection{Degenerate Situations}
\label{sec:degenerate}

Which two triangles, if intersecting, intersect on a line segment in most situations, they can also intersect on a point or a convex area (coplanar case). Even if line segment is the intersection, the intersection can be on triangle edges or inside triangles. These degenerate situations are troublesome to develop a robust boolean method. In this section, we offer simple but robust way to deal with all the degenerations and conceal the complexity of intersections behind the concise intersection representation in \S\ref{sec:ir}.

%The criterion of whether an intersection help tessellation is by its necessity: if the intersection is not presented, whether some faces from the linked halfedge will cross the boundary of primitives. The word 'cross' do not only include the situation that a face is partly outside and partly inside of a primitive, but also can mean that a face is partly inside (outside) and part on the boundary of the primitive.

\subsubsection{Point intersection}
\label{sec:ipoint}
If two triangles intersect on a single point (Fig. \ref{fig:isolated}), the intersection cannot be represented using our four-component description. Therefore, we simply add the intersection point into the meshes to guarantee correct tessellation, and no intersection is introduced.

@@

\subsubsection{Edge intersection}

% intersect on face edges does not create new line segments, it is part of the edges from other meshes, and will to be duplicated.

% whose edge? need to be further explained and better defined. an edge intersection divide edge in a triangle and will be duplicated in the other

When intersection line segment lies on an edge (Fig. \ref{fig:twin}). we call it \emph{edge intersection}. The only difference between edge intersection and the general case is the neighborhood representation $\Psi$. The space near edge intersection is decided by both face in the two side of the intersected edge, instead of just one triangle face (Fig. X). Therefore, supposing intersection is on the edge from mesh $P_k$, the neighborhood $\Psi_k$ is then an edge. Note that an intersection can be on edges from different meshes. Then there are more than one component of $\bm{\Psi}$ are edges.
